\ch{\textsc{RELIKE}}: Reionization Effective Likelihood from Planck 2018 Data (?)

### 2021-03-30 paper/rei.tex

\ch{I'm here}
Since the release of the Planck 2018 official analysis, an improved likelihood for the low-$\ell$ $E$-mode polarization was publicly released in 2019. This new likelihood,
$\texttt{srollv2}$\footnote{SRoll2: \url{http://sroll20.ias.u-psud.fr.}}
~\cite{Delouis:2019bub}, allows for improved reionization constraints with its better foreground modeling: The error bar on the optical depth $\tau$ in the tanh model is reduced by about $\sim40\%$ from XX to $\tau = 0.0566^{+0.0053}_{−0.0062}$~\cite{}.\ch{doesn't seem to be true anymore, 40\% was compared to the Planck 2018 likelihood paper result, the cosmo parameter paper seems to have similar error bars: $0.0519^{+0.0030}_{−0.0079}$.

\todo{CH: might need to change all instances of \textt{srollv2} to \textt{SRoll2} as used in the original sroll2 paper; also possibly \textt{simall\_EE} to \textt{SimAll} as used in the Planck 2018 paper.}

Since the Planck final release will likely be the best full-sky survey from space available to us in the next decade, it is important to extract all the information present in the data. In light of the $\texttt{srollv2}$ likelihood release, we obtain new reionization PC constraints with this latest likelihood. Enabled by the completeness property of the PCs, we also turn these constraints into an effective likelihood useful for assessing the CMB likelihood of \textit{any} reionization model out to $\zmax$ = 30, following techniques tested in Ref.~\cite{Heinrich:2016ojb}.

@@

The paper is structured as follows. We first describe in \S\ref{sec:background} the background on the reionization principal components and the kernel density estimate (KDE) technique used for building the effective likelihood. Then in \S\ref{sec:results}, we describe the PC results using the Planck 2018 + \texttt{srollv2} likelihood, and compare with previous results in literature. Then in \S\ref{sec:effective_likelihood}, we present the effective likelihood code, and demonstrate its fidelity with two examples, 1) the standard steplike model, and 2) a two-parameter model where a plateau of ionization at high-$z$ is added to the standard one-step model for illustration purposes. Finally, we summarize our results and conclude in \S\ref{sec:conclusion}.

@@

\section{Planck 2018 PC Results}
\label{sec:results}
We now present the complete reionization constraints obtained from the Planck 2018 likelihoods using the principal components.
\subsection{Constraints on Principal Components}
We use the official Planck likelihoods [add citation] \texttt{plik\_lite} for the high-$\ell$ $TT$, $TE$ and $EE$ as well as \texttt{lowl} for the low-$\ell$ $TT$ throughout this paper. We have tested that our results do not change if in lieu of \texttt{plik\_lite} we used the full \texttt{plik\_full} likelihood in which the foreground parameters have not been marginalized over. For the low-$\ell$ $EE$ likelihood, we use the third-party released \texttt{srollv2} likelihood [add citation] in our official PC results and the effective likelihood code. In comparison with the official Planck-released \texttt{simall\_EE} likelihood, the \texttt{srollv2} likelihood improves foreground cleaning, 
which enables stronger reionization constraints, especially for the low multipole moments that are important for the low redshift constraints.

@@

\begin{figure*}
%\includegraphics[width=0.7\textwidth]{plots/plot_mj_triangle_t18_r12_t19_t20_vs_pl18_pc_zmax30_pliklite_srollv2_1015_wTauTrajectory_pl15_wTanhML.png}
\includegraphics[width=0.8\textwidth]{paper/plots/plot_mj_triangle_t18_r12_t19_t20_vs_pl18_pc_zmax30_pliklite_srollv2_1015_wTauTrajectory_pl15_wTanhML_wGaussEllipse.pdf}
\caption{Constraints from Planck data on the amplitudes of the five reionization principal components that describe all physical ionization models up to $z_{\rm max} = 30$. We show  the 68\% and 95\% $m_a-m_b$ constraints as well as marginal constraints on individual $m_a$ for Planck 2018 in blue,
using \texttt{srollv2} for the low-$\ell$ $EE$ likelihood, and compare that to the Planck 2015 constraints in red.  The corresponding marginal distributions for individual $m_a$ are also shown.
%The 1D posterior distributions are shown on the diagonal and the 2D 68\% and 95\% C.L. regions in the $m_a-m_b$ planes in the lower triangle.
When deriving model independent optical depth constraints we impose an additional physicality prior that excludes the gray region in $m_1$ and regions beyond the box limits in the others.
 The black solid line indicates the trajectory of tanh models with varying $\tau$. The red and black crosses indicate the best-fit tanh models from Planck 2015 and 2018 respectively. Note that the best-fit tanh model has moved to a lower $\tau$ in Planck 2018, and in particular, is now within the 68\% C.L. contours of the constraints.
}
\label{fig:plot_mjs_2018_vs_2015}
\end{figure*}
%
%
%
%
In Fig.~\ref{fig:plot_mjs_2018_vs_2015}, we show the 1D posterior and 2D 68\% and 95\% confidence level contours for the 5 PC amplitudes that describe ionization models up to $\zmax=30$. We also show the trajectory of the tanh model (see \S \ref{sec:example1}) through this space as well as the best fit tanh model points.\footnote{The Planck 2015 best-fit for the tanh model is derived from a chain best-fit model as taken from Ref.~\cite{Heinrich:2016ojb}, whereas for the Planck 2018 point in this paper we used a minimizer to find the best-fit; there are also slight model differences, where in Planck 2018 we used a tanh width of $\Delta z = 0.015(1+z)$ instead of $\Delta z = 0.5$ for the Planck 2015 best-fit. Using $\Delta z = 0.5$ for Planck 2018 moves the cross negligibly on this plot.}
%
%
%
%
The gray shaded region % in the $m_1-m_2$ plane 
represents the parameter space that would be excluded by an additional 
physicality prior from Eq.~(\ref{eq:individualprior}),
which is just beyond the border of the displayed regime for all but $m_1$. Results using the Planck 2018 data are shown in blue. While all five PC amplitudes are  constrained by Planck, the first two are particularly well-constrained. Contrary to the Planck 2015 results of Ref.~\cite{Heinrich:2016ojb} shown in red, the 2018 data prefers a much smaller $m_1$, leading to a reduced optical depth, especially at high redshift.  
This preference also increases the impact of placing a physicality prior at low values of $m_1$ when deriving model-independent constraints as we shall discuss further in \S \ref{sec:modelindependent}.
We show a zoomed version of the $m_1-m_2$ plane in Fig.~\ref{fig:plot_m1m2_2015_vs_2018}.
In the physicality prior excluded region (gray), the ionization and the cumulative optical depth at high redshift becomes negative.

In Fig.~\ref{fig:plot_m1m2_2015_vs_2018}, we also show
the Planck 2018 constraints using the older \texttt{simall\_EE} likelihood.  The improvement 
from switching to the \texttt{srollv2} likelihood mainly strengthens the upper bound in the best constrained direction in the $m_1-m_2$ plane which shifts constraints toward higher total optical depth (see also Fig.~\ref{fig:prior_box}).

@@

\begin{figure}
\includegraphics[width=0.48\textwidth]{plots/plot_m1_m2_t18_r12_t19_t20_vs_pl18_pc_zmax30_pliklite_0930_vs_pl18_pc_zmax30_pliklite_srollv2_1015.pdf}
\caption{Changes in the best constrained $m_1-m_2$ PC plane between Planck 2015 (\wh{red}), Planck 2018 (blue, \texttt{srollv2}  \wh{decide whether to have this labeled}), and Planck 2018 but with the \texttt{simall\_EE} likelihood (\wh{green?}).   With Planck 2018, the \texttt{srollv2} likelihood moderately improves the best constrained direction and raises the optical depth.  }
\label{fig:plot_m1m2_2015_vs_2018}
\end{figure}

@@

%In Fig.~\ref{fig:plot_m1m2_lowE_vs_srollv2} we also compare the Planck 2018 results with two different low-$\ell$ $EE$ likelihoods: \texttt{simall\_EE} and \texttt{srollv2}. The \texttt{srollv2} contours are shrunk nearly directly along the lines of constant optical depth $\tau$ ($\tau_{12} = \tau_1 m_1 + \tau_2 m_2$) [wanna show this]? \wh{we might make this point in the context of Marius - that $\tau$ is mainly a linear function of $m_1$, $m_2$, and hence flat priors in $m's$ are locally flat in $\tau$}.

@@

\ch{Our code is available on GitHub at \url{...} [give URL]. } The initial release corresponds to the fiducial results in this paper, i.e.\ derived from the Planck 2018 $\texttt{plik\_lite\_TTTTTE} + \texttt{lowl} + \texttt{srollv2}$ likelihoods at $\zmax = 30$. 
The most important part of the code release is the python package called \texttt{Relike} which \wh{provides the PC parameters for a model and also} 
% added this since it now seemed to be emphasizing Gaussian only  - not sure that was intentional.
implements the Gaussian effective likelihood. The user can specify a customized $x_e(z)$ function taking in reionization model parameters and use \textt{Relike} to return the effective likelihood of the model. The returned likelihood is already marginalized over cosmological parameters. 
To obtain marginalized distributions more easily for high-dimensional parameter space, one may choose to sample \texttt{Relike} within a MCMC sampler such as Cobaya or Cosmosis \ch{[cite]}.

@@

Using Eq.~(\ref{eq:cumtauPC}), we can derive constraints on the total optical
depth $\tau_{\rm PC}=\tau_{\rm PC}(0,z_{\rm max})$ with the additional physicality priors on the PC amplitudes $m_a$ from
Eq.~\ref{eq:individualprior}.  These 
correspond to 
\beq
\tau_{\rm PC} = 0.0619^{+0.0056}_{-0.0068} \;\; (\texttt{srollv2}),
\label{eq:tau_pc_zmax30}
\eeq  
and the full posterior is shown in Fig.~\ref{fig:tauPC}. 
Notice that the distribution matches that of the two step model from \S \ref{sec:example2} but almost $0.5\sigma$ shifted higher than that of the tanh model from \S \ref{sec:example1}.

@@

Finally, we check that our results (Eq.~\ref{eq:tau_pc_zmax30}) are robust to
 extending the PC
range to
$z_{\rm max}=50$ which is negligibly different
\beq
\tau_{\rm PC,\, \zmax=50} = 0.0626^{+0.0061}_{-0.0072}\;\; (\texttt{srollv2}).
\eeq
Reverting the Planck 2018 likelihood to \texttt{simall\_EE} at $z_{\rm max}=30$ would give
\beq
\tau_{\rm PC} = 0.0582 ^{+0.0072}_{-0.0083}\;\; (\texttt{simall\_EE}).
\eeq
As with the tanh model, the \texttt{srollv2} likelihood provides tighter constraints that shift the distribution toward higher optical depth.

@@

In Fig.~\ref{fig:plot_taugtz_2018_with_vs_without_physicality_prior}, we show the 
the 68\% and 95\% confidence level contours for  $\tau_{\rm PC}(z,z_{\rm max})$ with and without the physicality prior on $m_1$.  Notice that without the prior, a larger range of unphysically negative optical depth is allowed whereas with the prior it is mostly eliminated except for the highest redshift.  Moreover upper bounds on the optical depth at high 
redshift are largely unchanged.   These upper limits allow some contribution at high-$z$ but do not require it.   For example
\beq
\tau_{\rm PC}(15, 30) < 0.023 \; (95\%\; \mathrm{C.L.})\;\;(\texttt{srollv2}).
\eeq
This should be contrasted with the tanh model 
 where it is  forbidden by form of the ionization history in conjunction with the total optical depth constraint.  
 On the other hand the bound is comparable to that obtained for the two-step model (see Eq.~\ref{eq:2stepbound}).  It is slightly larger given that the form of the two-step model still requires a steadily decreasing cumulative optical depth at $z>z_{\rm re}$).

@@

%Furthermore, their total optical depth constraints are practically identical: %
%\beq
%\tau_{\rm PC,\, \zmax=30} = 0.0619^{+0.0056}_{-0.0068}\;\; (\texttt{srollv2}),
%\eeq
%and
%\beq
%\tau_{\rm PC,\, \zmax=50} = 0.0626^{+0.0061}_{-0.0072}\;\; (\texttt{srollv2}).
%\eeq
%
%Compared the \texttt{simall\_EE} likelihood results, 
%\beq
%\tau_{\rm PC,\, \zmax=30} = 0.0582 ^{+0.0072}_{-0.0083}\;\; (\texttt{simall\_EE}),
%\eeq
%and
%\beq
%\tau_{\rm PC,\, \zmax=50} = 0.0582^{+0.0077}_{-0.0088}\;\; (\texttt{simall\_EE}).
%\eeq
%the \texttt{srollv2} error bars on PC $\tau$ was reduced by about 20\%, and the central values rose by about half a sigma. [wanna comment more here?]
%

For the high-redshift optical depth, we also obtain consistent upper limits
%\beq
%\tau_{\rm PC}(15, 30)< 0.023 \; (95\%\; %\mathrm{C.L.})\;\;(\texttt{srollv2}),
%\eeq
%[double check on full chain (from 10/15) this is from 09/30.]
%and 
\beq
\tau_{\rm PC}(15, 50) < 0.022 \; (95\%\; \mathrm{C.L.})\;\;(\texttt{srollv2}).
\eeq

@@

\begin{figure}[ht]
\includegraphics[width=0.45\textwidth]{paper/plots/pl18_taugtz_pc_zmax30_pl15_vs_pl18_simallEE_vs_pl18_srollv2_do_rescale_tau_false.pdf}
%previouslky
%{paper/plots/pl18_taugtz_pc_zmax30_pl15_vs_pl18_simallEE_vs_pl18_srollv2.pdf}
\caption{Changes in the
cumulative optical depth $\tau_{\rm PC}(z, \zmax)$ (as in Fig.~\ref{fig:plot_taugtz_2018_with_vs_without_physicality_prior}) 
between Planck 2015 and 2018 (top panel \wh{suggest swap to srollv2}) and between 2018 
\texttt{srollv2} (default) and $\texttt{simall\_EE}$.  Planck 2015 favored contributions at high $z$ which are disallowed in 2018 whereas the changes due to \texttt{srollv2} mainly tighten the lower limits at low $z$.
}
\label{fig:plot_taugtz_2015_vs_2018_simallEE_vs_2018_srollv2}
\end{figure}

\begin{figure}[ht]
\includegraphics[width=0.48\textwidth]{paper/plots/pl18_taugtz_pl18_srollv2_pc_zmax30_vs_zmax50_do_rescale_tau_false.pdf}
\caption{Robustness of the
cumulative optical depth $\tau_{\rm PC}(z, \zmax)$ to changes from the default $\zmax = 30$ to $\zmax = 50$. \wh{drop the Planck 2018 from label or at least make the $\zmax$ the prominent label}.  Upper limits at $z \lesssim 30$ are nearly unaffected whereas extending $\zmax$ allows the high-$z$ component to come from even higher-$z$.
}
\label{fig:plot_taugtz_zmax30_vs_zmax50}
\end{figure}

@@

In conclusion, we used constraints on reionization principal components with respect to the large-scale $EE$ data to produce a fast and accurate effective likelihood for the final release of Planck 2018 data. In lieu of the official $\texttt{simall\_EE}$ likelihood for low-$\ell$ $EE$ we used the independent later release of the $\texttt{srollv2}$ likelihood in 2019 which has better foreground cleaning allowing for slightly improved parameter constraints. Our effective likelihood code package is called $\textsc{Relike}$ [to be changed] and is available publicly on Github. This code is beneficial for the community for the purpose of testing \textit{any} global ionization history between $6 < z < 30$ with the Planck data without having to modify CAMB which can be very cumbersome to do. The convergence time to reach the Gilman-Rubin criteria $R_{-1} < 0.01$ is about 300 times less with the Relike code than with a traditional MCMC sampling of the Planck likelihood for the one-parameter canonical tanh model.

@@

    \item {The better foreground cleaned \texttt{srollv2} likelihood allowed to tighten the PC constraint on $\tau$ and shifted it toward slightly higher values when compared with the \texttt{plik\_lite\_TTTEEE+lowl+lowE} result  $\tau_{\rm PC} = 0.0582 ^{+0.0072}_{-0.0083}$ (68\% C.L.).}

>>>>

\ch{I'm here}
Since the release of the Planck 2018 official analysis, an improved likelihood for the low-$\ell$ $E$-mode polarization was publicly released in 2019. This new likelihood,
$\texttt{SRoll2}$\footnote{SRoll2: \url{http://sroll20.ias.u-psud.fr.}}
~\cite{Delouis:2019bub}, allows for improved reionization constraints with its better foreground modeling: The error bar on the optical depth $\tau$ in the tanh model is reduced by about $\sim40\%$ from XX to $\tau = 0.0566^{+0.0053}_{−0.0062}$~\cite{}.\ch{doesn't seem to be true anymore, 40\% was compared to the Planck 2018 likelihood paper result, the cosmo parameter paper seems to have similar error bars: $0.0519^{+0.0030}_{−0.0079}$.

Since the Planck final release will likely be the best full-sky survey from space available to us in the next decade, it is important to extract all the information present in the data. In light of the $\texttt{SRoll2}$ likelihood release, we obtain new reionization PC constraints with this latest likelihood. Enabled by the completeness property of the PCs, we also turn these constraints into an effective likelihood useful for assessing the CMB likelihood of \textit{any} reionization model out to $\zmax$ = 30, following techniques tested in Ref.~\cite{Heinrich:2016ojb}.

@@

The paper is structured as follows. We first describe in \S\ref{sec:background} the background on the reionization principal components and the kernel density estimate (KDE) technique used for building the effective likelihood. Then in \S\ref{sec:results}, we describe the PC results using the Planck 2018 + \texttt{SRoll2} likelihood, and compare with previous results in literature. Then in \S\ref{sec:effective_likelihood}, we present the effective likelihood code, and demonstrate its fidelity with two examples, 1) the standard steplike model, and 2) a two-parameter model where a plateau of ionization at high-$z$ is added to the standard one-step model for illustration purposes. Finally, we summarize our results and conclude in \S\ref{sec:conclusion}.

@@

\section{Planck 2018 PC Results}
\label{sec:results}
We now present the complete reionization constraints obtained from the Planck 2018 likelihoods using the principal components.
\subsection{Constraints on Principal Components}
We use the official Planck likelihoods [add citation] \texttt{plik\_lite} for the high-$\ell$ $TT$, $TE$ and $EE$ as well as \texttt{lowl} for the low-$\ell$ $TT$ throughout this paper. We have tested that our results do not change if in lieu of \texttt{plik\_lite} we used the full \texttt{plik\_full} likelihood in which the foreground parameters have not been marginalized over. For the low-$\ell$ $EE$ likelihood, we use the third-party released \texttt{SRoll2} likelihood [add citation] in our official PC results and the effective likelihood code. In comparison with the official Planck-released \texttt{SimAll} likelihood, the \texttt{SRoll2} likelihood improves foreground cleaning, 
which enables stronger reionization constraints, especially for the low multipole moments that are important for the low redshift constraints.

@@

\begin{figure*}
%\includegraphics[width=0.7\textwidth]{plots/plot_mj_triangle_t18_r12_t19_t20_vs_pl18_pc_zmax30_pliklite_srollv2_1015_wTauTrajectory_pl15_wTanhML.png}
\includegraphics[width=0.8\textwidth]{paper/plots/plot_mj_triangle_t18_r12_t19_t20_vs_pl18_pc_zmax30_pliklite_srollv2_1015_wTauTrajectory_pl15_wTanhML_wGaussEllipse.pdf}
\caption{Constraints from Planck data on the amplitudes of the five reionization principal components that describe all physical ionization models up to $z_{\rm max} = 30$. We show  the 68\% and 95\% $m_a-m_b$ constraints as well as marginal constraints on individual $m_a$ for Planck 2018 in blue,
using \texttt{SRoll2} for the low-$\ell$ $EE$ likelihood, and compare that to the Planck 2015 constraints in red.  The corresponding marginal distributions for individual $m_a$ are also shown.
%The 1D posterior distributions are shown on the diagonal and the 2D 68\% and 95\% C.L. regions in the $m_a-m_b$ planes in the lower triangle.
When deriving model independent optical depth constraints we impose an additional physicality prior that excludes the gray region in $m_1$ and regions beyond the box limits in the others.
 The black solid line indicates the trajectory of tanh models with varying $\tau$. The red and black crosses indicate the best-fit tanh models from Planck 2015 and 2018 respectively. Note that the best-fit tanh model has moved to a lower $\tau$ in Planck 2018, and in particular, is now within the 68\% C.L. contours of the constraints.
}
\label{fig:plot_mjs_2018_vs_2015}
\end{figure*}
%
%
%
%
In Fig.~\ref{fig:plot_mjs_2018_vs_2015}, we show the 1D posterior and 2D 68\% and 95\% confidence level contours for the 5 PC amplitudes that describe ionization models up to $\zmax=30$. We also show the trajectory of the tanh model (see \S \ref{sec:example1}) through this space as well as the best fit tanh model points.\footnote{The Planck 2015 best-fit for the tanh model is derived from a chain best-fit model as taken from Ref.~\cite{Heinrich:2016ojb}, whereas for the Planck 2018 point in this paper we used a minimizer to find the best-fit; there are also slight model differences, where in Planck 2018 we used a tanh width of $\Delta z = 0.015(1+z)$ instead of $\Delta z = 0.5$ for the Planck 2015 best-fit. Using $\Delta z = 0.5$ for Planck 2018 moves the cross negligibly on this plot.}
%
%
%
%
The gray shaded region % in the $m_1-m_2$ plane 
represents the parameter space that would be excluded by an additional 
physicality prior from Eq.~(\ref{eq:individualprior}),
which is just beyond the border of the displayed regime for all but $m_1$. Results using the Planck 2018 data are shown in blue. While all five PC amplitudes are  constrained by Planck, the first two are particularly well-constrained. Contrary to the Planck 2015 results of Ref.~\cite{Heinrich:2016ojb} shown in red, the 2018 data prefers a much smaller $m_1$, leading to a reduced optical depth, especially at high redshift.  
This preference also increases the impact of placing a physicality prior at low values of $m_1$ when deriving model-independent constraints as we shall discuss further in \S \ref{sec:modelindependent}.
We show a zoomed version of the $m_1-m_2$ plane in Fig.~\ref{fig:plot_m1m2_2015_vs_2018}.
In the physicality prior excluded region (gray), the ionization and the cumulative optical depth at high redshift becomes negative.

In Fig.~\ref{fig:plot_m1m2_2015_vs_2018}, we also show
the Planck 2018 constraints using the older \texttt{SimAll} likelihood.  The improvement 
from switching to the \texttt{SRoll2} likelihood mainly strengthens the upper bound in the best constrained direction in the $m_1-m_2$ plane which shifts constraints toward higher total optical depth (see also Fig.~\ref{fig:prior_box}).

@@

\begin{figure}
\includegraphics[width=0.48\textwidth]{plots/plot_m1_m2_t18_r12_t19_t20_vs_pl18_pc_zmax30_pliklite_0930_vs_pl18_pc_zmax30_pliklite_srollv2_1015.pdf}
\caption{Changes in the best constrained $m_1-m_2$ PC plane between Planck 2015 (\wh{red}), Planck 2018 (blue, \texttt{SRoll2}  \wh{decide whether to have this labeled}), and Planck 2018 but with the \texttt{SimAll} likelihood (\wh{green?}).   With Planck 2018, the \texttt{SRoll2} likelihood moderately improves the best constrained direction and raises the optical depth.  }
\label{fig:plot_m1m2_2015_vs_2018}
\end{figure}

@@

%In Fig.~\ref{fig:plot_m1m2_lowE_vs_srollv2} we also compare the Planck 2018 results with two different low-$\ell$ $EE$ likelihoods: \texttt{SimAll} and \texttt{SRoll2}. The \texttt{SRoll2} contours are shrunk nearly directly along the lines of constant optical depth $\tau$ ($\tau_{12} = \tau_1 m_1 + \tau_2 m_2$) [wanna show this]? \wh{we might make this point in the context of Marius - that $\tau$ is mainly a linear function of $m_1$, $m_2$, and hence flat priors in $m's$ are locally flat in $\tau$}.

@@

\ch{Our code is available on GitHub at \url{...} [give URL]. } The initial release corresponds to the fiducial results in this paper, i.e.\ derived from the Planck 2018 $\texttt{plik\_lite\_TTTTTE} + \texttt{lowl} + \texttt{SRoll2}$ likelihoods at $\zmax = 30$. 
The most important part of the code release is the python package called \texttt{Relike} which \wh{provides the PC parameters for a model and also} 
% added this since it now seemed to be emphasizing Gaussian only  - not sure that was intentional.
implements the Gaussian effective likelihood. The user can specify a customized $x_e(z)$ function taking in reionization model parameters and use \textt{Relike} to return the effective likelihood of the model. The returned likelihood is already marginalized over cosmological parameters. 
To obtain marginalized distributions more easily for high-dimensional parameter space, one may choose to sample \texttt{Relike} within a MCMC sampler such as Cobaya or Cosmosis \ch{[cite]}.

@@

Using Eq.~(\ref{eq:cumtauPC}), we can derive constraints on the total optical
depth $\tau_{\rm PC}=\tau_{\rm PC}(0,z_{\rm max})$ with the additional physicality priors on the PC amplitudes $m_a$ from
Eq.~\ref{eq:individualprior}.  These 
correspond to 
\beq
\tau_{\rm PC} = 0.0619^{+0.0056}_{-0.0068} \;\; (\texttt{SRoll2}),
\label{eq:tau_pc_zmax30}
\eeq  
and the full posterior is shown in Fig.~\ref{fig:tauPC}. 
Notice that the distribution matches that of the two step model from \S \ref{sec:example2} but almost $0.5\sigma$ shifted higher than that of the tanh model from \S \ref{sec:example1}.

@@

Finally, we check that our results (Eq.~\ref{eq:tau_pc_zmax30}) are robust to
 extending the PC
range to
$z_{\rm max}=50$ which is negligibly different
\beq
\tau_{\rm PC,\, \zmax=50} = 0.0626^{+0.0061}_{-0.0072}\;\; (\texttt{SRoll2}).
\eeq
Reverting the Planck 2018 likelihood to \texttt{SimAll} at $z_{\rm max}=30$ would give
\beq
\tau_{\rm PC} = 0.0582 ^{+0.0072}_{-0.0083}\;\; (\texttt{SimAll}).
\eeq
As with the tanh model, the \texttt{SRoll2} likelihood provides tighter constraints that shift the distribution toward higher optical depth.

@@

In Fig.~\ref{fig:plot_taugtz_2018_with_vs_without_physicality_prior}, we show the 
the 68\% and 95\% confidence level contours for  $\tau_{\rm PC}(z,z_{\rm max})$ with and without the physicality prior on $m_1$.  Notice that without the prior, a larger range of unphysically negative optical depth is allowed whereas with the prior it is mostly eliminated except for the highest redshift.  Moreover upper bounds on the optical depth at high 
redshift are largely unchanged.   These upper limits allow some contribution at high-$z$ but do not require it.   For example
\beq
\tau_{\rm PC}(15, 30) < 0.023 \; (95\%\; \mathrm{C.L.})\;\;(\texttt{SRoll2}).
\eeq
This should be contrasted with the tanh model 
 where it is  forbidden by form of the ionization history in conjunction with the total optical depth constraint.  
 On the other hand the bound is comparable to that obtained for the two-step model (see Eq.~\ref{eq:2stepbound}).  It is slightly larger given that the form of the two-step model still requires a steadily decreasing cumulative optical depth at $z>z_{\rm re}$).

@@

%Furthermore, their total optical depth constraints are practically identical: %
%\beq
%\tau_{\rm PC,\, \zmax=30} = 0.0619^{+0.0056}_{-0.0068}\;\; (\texttt{SRoll2}),
%\eeq
%and
%\beq
%\tau_{\rm PC,\, \zmax=50} = 0.0626^{+0.0061}_{-0.0072}\;\; (\texttt{SRoll2}).
%\eeq
%
%Compared the \texttt{SimAll} likelihood results, 
%\beq
%\tau_{\rm PC,\, \zmax=30} = 0.0582 ^{+0.0072}_{-0.0083}\;\; (\texttt{SimAll}),
%\eeq
%and
%\beq
%\tau_{\rm PC,\, \zmax=50} = 0.0582^{+0.0077}_{-0.0088}\;\; (\texttt{SimAll}).
%\eeq
%the \texttt{SRoll2} error bars on PC $\tau$ was reduced by about 20\%, and the central values rose by about half a sigma. [wanna comment more here?]
%

For the high-redshift optical depth, we also obtain consistent upper limits
%\beq
%\tau_{\rm PC}(15, 30)< 0.023 \; (95\%\; %\mathrm{C.L.})\;\;(\texttt{SRoll2}),
%\eeq
%[double check on full chain (from 10/15) this is from 09/30.]
%and 
\beq
\tau_{\rm PC}(15, 50) < 0.022 \; (95\%\; \mathrm{C.L.})\;\;(\texttt{SRoll2}).
\eeq

@@

\begin{figure}[ht]
\includegraphics[width=0.45\textwidth]{paper/plots/pl18_taugtz_pc_zmax30_pl15_vs_pl18_simallEE_vs_pl18_srollv2_do_rescale_tau_false.pdf}
%previouslky
%{paper/plots/pl18_taugtz_pc_zmax30_pl15_vs_pl18_simallEE_vs_pl18_srollv2.pdf}
\caption{Changes in the
cumulative optical depth $\tau_{\rm PC}(z, \zmax)$ (as in Fig.~\ref{fig:plot_taugtz_2018_with_vs_without_physicality_prior}) 
between Planck 2015 and 2018 (top panel) and between 2018 \texttt{SRoll2} (default) and $\texttt{SimAll}$.  Planck 2015 favored contributions at high $z$ which are disallowed in 2018 whereas the changes due to \texttt{SRoll2} mainly tighten the lower limits at low $z$.
}
\label{fig:plot_taugtz_2015_vs_2018_simallEE_vs_2018_srollv2}
\end{figure}

\begin{figure}[ht]
\includegraphics[width=0.48\textwidth]{paper/plots/pl18_taugtz_pl18_srollv2_pc_zmax30_vs_zmax50_do_rescale_tau_false.pdf}
\caption{Robustness of the \ch{Planck 2018}
cumulative optical depth $\tau_{\rm PC}(z, \zmax)$ to changes from the default $\zmax = 30$ to $\zmax = 50$. Upper limits at $z \lesssim 30$ are nearly unaffected whereas extending $\zmax$ allows the high-$z$ component to come from even higher-$z$.
}
\label{fig:plot_taugtz_zmax30_vs_zmax50}
\end{figure}

@@

In conclusion, we used constraints on reionization principal components with respect to the large-scale $EE$ data to produce a fast and accurate effective likelihood for the final release of Planck 2018 data. In lieu of the official $\texttt{SimAll}$ likelihood for low-$\ell$ $EE$ we used the independent later release of the $\texttt{SRoll2}$ likelihood in 2019 which has better foreground cleaning allowing for slightly improved parameter constraints. Our effective likelihood code package is called $\textsc{Relike}$ [to be changed] and is available publicly on Github. This code is beneficial for the community for the purpose of testing \textit{any} global ionization history between $6 < z < 30$ with the Planck data without having to modify CAMB which can be very cumbersome to do. The convergence time to reach the Gilman-Rubin criteria $R_{-1} < 0.01$ is about 300 times less with the Relike code than with a traditional MCMC sampling of the Planck likelihood for the one-parameter canonical tanh model.

@@

    \item {The better foreground cleaned \texttt{SRoll2} likelihood allowed to tighten the PC constraint on $\tau$ and shifted it toward slightly higher values when compared with the \texttt{plik\_lite\_TTTEEE+lowl+lowE} result  $\tau_{\rm PC} = 0.0582 ^{+0.0072}_{-0.0083}$ (68\% C.L.).}
